Kids First or Marriage First?

### 2019-05-31 main/draft.tex
Received fixes from Martí

\begin{abstract}
Unplanned pregnancies often trigger marriage for unmarried cohabitating couples. I show that on average marriages in which a child appeared before or at the year of marriage are characterized with higher divorce probabilities and other outcomes suggesting lower marital quality. This association is robust to different ways of controlling for composition, and it is more pronounced for higher income and more educated couples. I argue that this pattern is consistent with household bargaining models, and the main driver of the result is unequal exposure of spouses to the childcare costs: since being single mother is a bad outcome for an educated female, she is more likely to agree to enter the marriages of lower quality. I estimate a lifecycle model with dynamic bargaining and fertility decisions to show how much of these differences is driven by household bargaining and how unplanned pregnancies creating unexpected childcare costs contribute to observable divorce rates.
\end{abstract}

@@

\section{Reduced Form Evidence}
\subsection{Raw Comparisons}
First, I compare divorce probabilities in cross-sections of people between $K\to M$ and $M\to K$ groups. See Table \ref{diff-raw} for this.

\begin{table}[h!]
\begin{center}
\begin{tabular}{|l|c||c|c||c||c|c|}\cline{2-7}
\multicolumn{1}{c|}{} & \multicolumn{3}{c||}{\small ACS: sample of married once} & \multicolumn{3}{c|}{\small SIPP: sample of  ever married} \\\cline{2-7}
\multicolumn{1}{c|}{}  & \multirow{2}{*}{\small \% K$\to$M}& \multicolumn{2}{c||}{\textbf{\% divorced (now)}} & \multirow{2}{*}{\footnotesize \% K$\to$M }& \multicolumn{2}{c|}{\textbf{\% divorced (ever)}} \\\cline{3-4}\cline{6-7}
\multicolumn{1}{c|}{}  &  &\small if M$\to$K &  \small if K$\to$M & & \small if M$\to$K & \small if K$\to$M \\\hline
\textit{All females, 21--40}  & $23.2\%$ & $10.1\%$ & $18.0\%$ & $22.8\%$ & $24.1\%$ & $34.8\%$ \\
\textit{All females, 35--40} & $18.0\%$ & $11.3\%$ & $23.0\%$ & $18.9\%$ & $28.1\%$ & $45.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 21--40} & $34.2\%$ & $13.9\%$ & $17.2\%$ & $33.4\%$ & $29.2\%$  & $34.2\%$ \\
\textit{College graduates, 21--40} & $9.9\%$ & $5.3\%$ & $14.7\%$ & $10.5\%$ & $12.9\%$ & $27.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} per-hour income for subsample of working women.}\\\hline
\textit{Low 30\% income, 21--40} & $30.8\%$ & $17.6\%$ & $24.1\%$ & $34.6\%$ & $30.5\%$ &  $42.5\%$ \\
\textit{Mid 40\% income, 21--40} & $25.5\%$ & $14.2\%$ & $21.5\%$ & $24.1\%$ & $28.3\%$ & $45.8\%$ \\
\textit{Top 30\% income, 21--40} & $14.1\%$ & $7.7\%$ & $18.1\%$ & $12.4\%$ & $22.9\%$  & $32.8\%$ \\\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize $\Delta T = \text{Year}^{\text{first child}} - \text{Year}^{\text{first marriage}}$. Restriction: $\Delta T \in \{-5,...,10\}$.}\\
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize Definition: K$\to$M if $\Delta T \in \{-5,...,0\}$. M$\to$K if $\Delta T \in \{1,...,10\}$.}\\
\hline
\end{tabular}
\caption{Differences in raw means, ACS and SIPP\label{diff-raw}}
\end{center}
\end{table}

@@

\subsection{Differences in Observables}
This section captures demographic and economic differences of K$\to$M and M$\to$K groups. Many factors could compute to the observable differences, here I quantify them and this justifies use of controls in the next part. I present these differences for ACS, SIPP tables are available in [Appendix ...]. For ACS, since M$\to$K and K$\to$M are defined only for subsample of people who were married once and have children present, I also compare this subsample with more general female population. See Table \ref{diff-contr} for the results.

@@

\subsection{Overview}
Full list of parameters involves many things:
\begin{itemize}
\item Fundamental parameters of income fluctuation model $(\beta,\sigma,R)$
\item Intra-household returns to scale $\rho_c$
\item Preferences for child's quality: $(\phi_0,\phi_1,\phi_2)$
\item Parameters for child's quality production function: CES power $\theta_k$ and labor shares by age $(\mu_y,\mu_s)$.
\item Transition probabilities:
\begin{itemize}
\item Probability of pregnancy shock by age: $p^{\text{preg}}_t = \lambda^{\text{preg}}_0 - \lambda^{\text{preg}}_1\cdot t$ (2 parameters)
\item Probability of meeting a partner: $p^{\text{meet}}_t:$ piecewise linear function between $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$ (4 parameters)
\item Probability of re-entering marriage market and corresponding utility ``compensation'' $(p^{\text{out}},S)$. (2 parameters)
\item Probability that kids grow up (transition from $y$ to $s$): $p_s$ (1 parameter)
\end{itemize}
\item Age profiles of labor income by gender: $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ (6 parameters)
\item Variance of shocks (6 parameters):
\begin{itemize}
\item Individual income shocks $\sigma^{z,f}$, $\sigma^{z,m}$.
\item Couple's surplus shocks $\sigma^{\psi}$
\item Couple's initial surplus $\sigma^{\psi_0}$
\item Assets and productivity of potential partner: $(\sigma^{z,p},\sigma^{a,p})$
\end{itemize}
\end{itemize}

Total number of parameters here is 31. Few additional choices involve time horizons, bounds for random walk variables and bounds for assets and $\theta$.

\subsection{Externally Set Parameters}
Length of the period is set to be 1 year, the model assumes people start their life with no assets at the age of $21$ and die deterministically at the age of $81$ (so $T = 30$). However, it assumes that fertility status can change only till age of 40 ($T_f = 20$) and marriage status till age of 45 ($T_m = 25$). It also assumes that after the age of 60 people retire and get $b = 0.4$ [as in Sommers, 2014] of their income at age of 60 (so no transitions happen). I only simulate the data for age 21--40, therefore these choices are not likely to play a crucial role.

Following standard approach in literature, I fix $\sigma = 1.5$, $R = 1.03$ and $\beta = 1/R$. %Also, following estimates of Voena (2014) I set $\rho_c = 0.23$.

I take age-income profiles  $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ to match coefficients in per-hour labor earnings in ACS (see Appendix [...] for details). I borrow variances of income shocks $\sigma^{z,f}$, $\sigma^{z,m}$ from Low et al 2018 as they cannot be estimated in cross-sectional data [This could be fixed in later versions as their variances are for subsample of non-college graduates].

@@

In the part with assets, I assume savings can be between 0 and 20 times average labor income in the beginning of life. This bar in pretty low, but it is enough to create reasonable assets variation at the moment most of marriage decisions are made: say, median assets-to-income ratio at 30 is around 1 [see ...].

@@

Conceptually, they can be split onto following groups:
\begin{enumerate}
\item Unplanned pregnancy probabilities are pinned down by share of $K\to M$ females at ages 25 and 35, as well as share of planned births in total population
\item Meeting probabilities and $\sigma^{\psi_0}$ are pinned down by shares of never married females people at 25, 35, 40 and share of females with one marriage at 40 (relevant to capture chances of re-marrying).
\item  Preferences for kids are captured by share of childless females by age of 25, 30, 35 (identifies general scale $\phi_0$), share of childless females for low and high income at 30 $\phi_1$. Luxury good parameter $\phi_2$ is pinned down by ratio of amounts of money spent on kids between 75 and 25 income quantiles.
\item Childcare production function is recovered by ratio of time spent with kids by couple's income (pins down elasticity of substitution), average share of money spent on kids (captures ``average'' $\mu$) and ratio of time spent by child's age (captures difference between $\mu_a$ and $\mu_y$)
\item Variance of innovation to marriage surplus is mainly pinned down by share of divorced people by age (25, 35, 40).
\item Transition out of single mothers is captured by share of never married single mothers at 25 (picks $S$) and share of divorced women with kids at 30 (picks $p^{\text{out}}$).
\item Returns to scale in marriage $\rho_c$ are pinned down by comparing divorce probabilities between high and low income females.
\end{enumerate}
On top of it, I use shares of divorced people for $K\to M$ and $M \to K$ at 40 and their ratio by income groups as additional targets that are affected by mixture of parameters described above.

@@

\section{Results}
Table \ref{targets-main} shows the targets in the model and in the data. [Marriage dimension is somewhat off, this is not crucial as the study mainly contains married couples, though I have few ideas how to fix it in finite time.]
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|}\hline
Target & Model & Data \\\hline
\multicolumn{3}{|c|}{Marriage Dimension} \\\hline
Never married by 25  & $32.8\%$ & $68.3\%$  \\
Never married by 30 & $12.6\%$ & $25.6\%$  \\
Never married by 40 & $6.3\%$ & $18.8\%$  \\\hline
One marriage by 40 & $66.3\%$ & $64.7\%$ \\\hline
\multicolumn{3}{|c|}{Fertility Dimension} \\\hline
No kids at 25 & $59.0\%$ & $67.7\%$ \\
No kids at 30 (low $30\%$ income) & $46.7\%$  &   $44.0\%$ \\
No kids at 30 (top $30\%$ income) & $52.1\%$  &   $57.8\%$ \\
No kids at 35  & $32.9\%$  &   $28.1\%$ \\
No kids at 40 & $17.6\%$ & $25.6\%$ \\\hline
\multicolumn{3}{|c|}{Marriage/Fertility Interactions} \\\hline
Never married with kids at 30 & $7.0\%$  & $12.3\%$ \\
Divorced with kids at 30 & $5.7\%$ & $3.8\%$ \\\hline
\multicolumn{3}{|c|}{Divorce Dimension (Divorced = Divorced and not remarried)} \\\hline
Divorced by 25 & 6.6\%   &  10.7\% \\
Divorced by 30 (low $30\%$ income)  & 14.5\% &     20.8\% \\
Divorced by 30 (top $30\%$ income)  & 14.0\% &     8.1\% \\
Divorced by 35  & 15.8\% &    15.9\% \\
Divorced by 40  & 15.5\% &    19.1\% \\\hline
\multicolumn{3}{|c|}{K$\to$M --- M$\to$K partition} \\\hline
Divorced by 40 if K$\to$M & 22.1\% &   23.5\% \\
Divorced by 40 if M$\to$K & 13.8\% &      13.0\% \\
Divorced by 40 if K$\to$M and low 30\% & 34.0\% & 28.5\% \\
Divorced by 40 if K$\to$M and top 30\% & 28.9\% & 23.3\% \\
Divorced by 40 if M$\to$K and low 30\% & 34.0\% & 28.5\% \\
Divorced by 40 if M$\to$K and top 30\% & 19.3\% & 11.3\% \\
Divorced by 40 if M$\to$K and low 30\% & 28.7\% & 18.1\% \\\hline
\multicolumn{3}{|c|}{Childcare} \\\hline
Money for childcare / labor income & 25.7\% & 40.0\% \\
Time spent 0--5 / Time spent 6--12 & 1.92 & 2 \\
Time spent top 30\% / Time spent low 30\% & 0.92 & 1.19 \\
Money spent top 30\% / Money spent low 30\% & 2.09 & 2.67 \\\hline

\end{tabular}
\caption{Targets in the data and in the model\label{targets-main}}
\end{center}
\end{table}

\subsection{Counterfactual Simulations}
[TBD]
\begin{itemize}
\item Turn off fertility completely
\item Turn off bargaining
\item No pressure of being single mother
\item Single mother/single father randomly
\end{itemize}

\subsubsection{Altering Fertility Shocks}
I consider two counterfactual scenarios: when unexpected fertility never happens at the moment people meet and when it never happens at all (but people still can make kids by choice). I compare key targets in Table \ref{cf-shocks}. In general, it looks like contribution of fertility shocks to the divorce rates is quite substantial, and major part of it is driven by fertility shocks before the marriage happens. Especially reducing fertility shocks affects divorces in early stages.
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|}\hline
Target & Baseline & No shocks at meeting & No shocks at all \\\hline
\multicolumn{4}{|c|}{Marriage Dimension} \\\hline
Never married by 25  & $32.8\%$ & $30.6\%$ & $31.1\%$  \\
Never married by 30 & $12.6\%$ & $10.8\%$ & $11.2\%$  \\
Never married by 40 & $6.3\%$ & $5.7\%$ & $6.1\%$  \\\hline
\multicolumn{4}{|c|}{Fertility Dimension} \\\hline
No kids at 25 & $59.0\%$ & $66.3\%$ & $68.3\%$ \\
No kids at 30 (low $30\%$ income) & $46.7\%$  &   $53.5\%$ & $54.8\%$ \\
No kids at 30 (top $30\%$ income) & $52.1\%$  &   $61.9\%$ & $62.7\%$ \\
No kids at 35  & $32.9\%$  &   $36.2\%$ & $36.6\%$ \\\hline
\multicolumn{4}{|c|}{Divorce Dimension} \\\hline
Divorced by 25 & 6.6\%   &  2.6\% & 1.4\% \\
Divorced by 30 (low $30\%$ income)  & 14.5\% &     11.0\% & 8.6\% \\
Divorced by 30 (top $30\%$ income)  & 14.0\% &     9.0\% & 7.5\% \\
Divorced by 35  & 15.8\% &    13.6\% & 11.9\% \\
Divorced by 40  & 15.5\% &    14.3\% & 13.0\% \\\hline
\end{tabular}
\caption{Contribution of fertility shocks \label{cf-shocks}}
\end{center}
\end{table}

\section{Robustness Checks}
\section{Conclusions}
\clearpage
\section*{Literature}
\bibliography{bib}

>>>>

\begin{abstract}
Unplanned pregnancies often trigger marriage for unmarried couples. I show that, on average, marriages in which a child appeared before or at the year of marriage are characterized with higher divorce probabilities and other outcomes suggesting lower marital quality. This association is robust to different ways of controlling for composition, and it is more pronounced for higher income and more educated couples. I argue that this pattern is consistent with household bargaining models, and the main driver of the result is unequal exposure of spouses to the childcare costs: since being single mother is a bad outcome for an educated female, she is more likely to agree to enter the marriages of lower quality. I estimate a lifecycle model with dynamic bargaining and fertility decisions to show how much of these differences is driven by household bargaining and how unplanned pregnancies creating unexpected childcare costs contribute to observable divorce rates.
\end{abstract}

@@

\section{Reduced Form Evidence}
\subsection{Raw Comparisons}
The main result I discuss is summarized by Figure \ref{shrs}. Table \ref{diff-raw} presents these results in numbers.

\begin{figure}[h!]
\includegraphics[width=0.9\linewidth]{shrs.eps}
\caption{Share of K$\to$M couples and share of divorced\label{shrs}}
\end{figure}

%First, I compare divorce probabilities in cross-sections of people between $K\to M$ and $M\to K$ groups. See Table \ref{diff-raw} for this.

\begin{table}[p]
\begin{center}
\begin{tabular}{|l|c||c|c||c||c|c|}\cline{2-7}
\multicolumn{1}{c|}{} & \multicolumn{3}{c||}{\small ACS: sample of married once} & \multicolumn{3}{c|}{\small SIPP: sample of  ever married} \\\cline{2-7}
\multicolumn{1}{c|}{}  & \multirow{2}{*}{\small \% K$\to$M}& \multicolumn{2}{c||}{\textbf{\% divorced (now)}} & \multirow{2}{*}{\footnotesize \% K$\to$M }& \multicolumn{2}{c|}{\textbf{\% divorced (ever)}} \\\cline{3-4}\cline{6-7}
\multicolumn{1}{c|}{}  &  &\small if M$\to$K &  \small if K$\to$M & & \small if M$\to$K & \small if K$\to$M \\\hline
\textit{All females, 21--40}  & $23.2\%$ & $10.1\%$ & $18.0\%$ & $22.8\%$ & $24.1\%$ & $34.8\%$ \\
\textit{All females, 35--40} & $18.0\%$ & $11.3\%$ & $23.0\%$ & $18.9\%$ & $28.1\%$ & $45.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 21--40} & $34.2\%$ & $13.9\%$ & $17.2\%$ & $33.4\%$ & $29.2\%$  & $34.2\%$ \\
\textit{College graduates, 21--40} & $9.9\%$ & $5.3\%$ & $14.7\%$ & $10.5\%$ & $12.9\%$ & $27.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} per-hour income for subsample of working women.}\\\hline
\textit{Low 30\% income, 21--40} & $30.8\%$ & $17.6\%$ & $24.1\%$ & $34.6\%$ & $30.5\%$ &  $42.5\%$ \\
\textit{Mid 40\% income, 21--40} & $25.5\%$ & $14.2\%$ & $21.5\%$ & $24.1\%$ & $28.3\%$ & $45.8\%$ \\
\textit{Top 30\% income, 21--40} & $14.1\%$ & $7.7\%$ & $18.1\%$ & $12.4\%$ & $22.9\%$  & $32.8\%$ \\\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize $\Delta T = \text{Year}^{\text{first child}} - \text{Year}^{\text{first marriage}}$. Restriction: $\Delta T \in \{-5,...,10\}$.}\\
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize Definition: K$\to$M if $\Delta T \in \{-5,...,0\}$. M$\to$K if $\Delta T \in \{1,...,10\}$.}\\
\hline
\end{tabular}
\caption{Differences in raw means, ACS and SIPP\label{diff-raw}}
\end{center}
\end{table}

@@

\subsection{Differences in Observables}
This section captures demographic and economic differences of K$\to$M and M$\to$K groups. Many factors could compute to the observable differences, here I quantify them and this justifies use of controls in the next part. I present these differences for ACS, SIPP data are not very different except smaller sample size. For ACS, since M$\to$K and K$\to$M are defined only for subsample of people who were married once and have children present, I also compare this subsample with more general female population. See Table \ref{diff-contr} for the results.

@@

\subsection{Overview}
Full list of parameters involves many things:
\begin{itemize}
\item Fundamental parameters of income fluctuation model $(\beta,\sigma,R)$
\item Preferences for child's quality: $(\phi_0,\phi_1,\phi_2,\phi_3)$
\item Parameters for child's quality production function: CES power $\theta_k$ and labor shares by age $(\mu_y,\mu_s)$
\item Unavoidable childcare costs $\underline{q}$
\item Transition probabilities:
\begin{itemize}
\item Probability of pregnancy shock by age: $p^{\text{preg}}_t = \lambda^{\text{preg}}_0 - \lambda^{\text{preg}}_1\cdot t$ (2 parameters)
\item Probability of meeting a partner: $p^{\text{meet}}_t:$ piecewise linear function between $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$ (4 parameters)
\item Probability of re-entering marriage market and corresponding utility ``compensation'' $(p^{\text{out}},S)$. I set $S = 0$ in current calibration (so 1 parameter).
\item Probability that kids grow up (transition from $y$ to $s$): $p_s$ (1 parameter)
\end{itemize}
\item Age profiles of labor income by gender: $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ (6 parameters)
\item Variance of shocks (6 parameters):
\begin{itemize}
\item Individual income shocks $\sigma^{z,f}$, $\sigma^{z,m}$.
\item Couple's surplus shocks $\sigma^{\psi}$
\item Couple's initial surplus $\sigma^{\psi_0}$
\item Assets and productivity of potential partner: $(\sigma^{z,p},\sigma^{a,p})$
\end{itemize}
\end{itemize}

Few additional choices involve time horizons, bounds for random walk variables and bounds for assets and $\theta$.

\subsection{Externally Set Parameters}
Length of the period is set to be 1 year, the model assumes people start their life with no assets at the age of $21$ and die deterministically at the age of $81$ (so $T = 60$). However, it assumes that fertility status can change only till age of 37 ($T_f = 17$) and marriage status till age of 45 ($T_m = 25$). It also assumes that after the age of 60 people retire and get $b = 0.4$ as in Sommer, 2014 of their income at age of 60 (so no transitions happen after $T = 40$). I only simulate the data for age 21--40, therefore these choices are not likely to play a crucial role.

Following standard approach in the macro literature, I fix $\sigma = 2$, $R = 1.03$ and $\beta = 1/R$. Also, following estimates of Voena, 2014 I set $\rho_c = 0.23$.

I take age-income profiles  $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ to match coefficients in per-hour labor earnings in ACS (see Appendix [...] for details). I borrow variances of income shocks $\sigma^{z,f}$, $\sigma^{z,m}$ from Low et al 2018 as they cannot be estimated in cross-sectional data, and although they were estimated for subsample of non college graduates in SIPP, in the model they produce IQR at the age of 30 that is very similar to the full sample of ACS data.

@@

In the part with assets, I assume savings can be between 0 and 20 times average labor income in the beginning of life. This bar in pretty low, but it is enough to create reasonable assets variation at the moment most of marriage decisions are made: say, median assets-to-income ratio at 30 is around 1.

@@

Conceptually, they can be split onto following groups:
\begin{enumerate}
\item Unplanned pregnancy probabilities are pinned down by share of $K\to M$ females at ages 25 and 35, as well as share of planned births in total population
\item Meeting probabilities and $\sigma^{\psi_0}$ are pinned down by shares of never married females people at 25, 35, 40 and share of females with one marriage at 40 (relevant to capture chances of re-marrying).
\item  Preferences for kids are captured by share of childless females by age of 25, 30, 35 (identifies general scale $\phi_0$ and $\phi_3$), share of childless females for low and high income at 30 (captures $\phi_0$ and $\phi_1$). Luxury good parameter $\phi_2$ is mainly pinned down by ratio of amounts of money spent on kids between 75 and 25 income quantiles.
\item Childcare production function is recovered by ratio of time spent with kids by couple's income (pins down elasticity of substitution), average share of money spent on kids (captures ``average'' $\mu$) and ratio of time spent by child's age (captures difference between $\mu_a$ and $\mu_y$)
\item Variance of innovation to marriage surplus is mainly pinned down by share of divorced people by age (25, 35, 40) and divorce rates by income groups.
\item Transition out of single mothers is captured by share of never married single mothers at 25 and share of divorced women with kids at 30 (picks $p^{\text{out}}$).
\end{enumerate}
On top of it, I use shares of divorced people for $K\to M$ and $M \to K$ at 40 per se and by income groups as additional targets that are affected by mixture of parameters described above.

@@

\section{Results}
Table \ref{targets-main} shows the targets in the model and in the data. [Income partition is somewhat off, but things get better over time.]

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|}\hline
Target & Model & Data \\\hline
\multicolumn{3}{|c|}{Marriage Dimension} \\\hline
Never married by 25  & $67.8\%$ & $68.3\%$ \\
Never married by 30 & $25.2\%$ & $25.6\%$  \\
Never married by 40 & $20.6\%$ & $18.8\%$  \\\hline
One marriage by 40  & $68.9\%$ & $64.7\%$ \\\hline
\multicolumn{3}{|c|}{Fertility Dimension} \\\hline
No kids at 25 & $71.7\%$ & $67.7\%$ \\
No kids at 30 (low $30\%$ income) & $58.6\%$  &   $44.0\%$ \\
No kids at 30 (top $30\%$ income) & $56.6\%$  &   $57.8\%$ \\
No kids at 35  & $22.4\%$  &   $28.1\%$ \\
No kids at 40 & $19.4\%$ & $25.6\%$ \\\hline
\multicolumn{3}{|c|}{Marriage/Fertility Interactions} \\\hline
Never married with kids at 30 & $12.4\%$  & $12.3\%$ \\
Divorced with kids at 30 & $3.2\%$ & $3.8\%$ \\\hline
\multicolumn{3}{|c|}{Divorce Dimension (Divorced $=$ Divorced and not remarried)} \\\hline
Divorced by 25 & $4.9\%$   & $10.7\%$ \\
Divorced by 30 (low $30\%$ income)  & 15.6\% &     20.8\% \\
Divorced by 30 (top $30\%$ income)  & 9.9\% &     8.1\% \\
Divorced by 35  & 8.4\% &    15.9\% \\
Divorced by 40  & 10.6\% &    19.1\% \\\hline
\multicolumn{3}{|c|}{K$\to$M --- M$\to$K partition} \\\hline
Divorced by 40 if K$\to$M & 20.2\% &   23.5\% \\
Divorced by 40 if M$\to$K & 8.7\% &      13.0\% \\
Divorced by 40 if K$\to$M and low 30\% & 34.0\% & 28.5\% \\
Divorced by 40 if K$\to$M and top 30\% & 26.6\% & 23.3\% \\
Divorced by 40 if M$\to$K and low 30\% & 15.3\% & 18.0\% \\
Divorced by 40 if M$\to$K and top 30\% & 9.3\% & 11.3\% \\\hline
\multicolumn{3}{|c|}{Childcare} \\\hline
Money for childcare / labor income & 40.1\% & 40.0\% \\
Time spent 0--5 / Time spent 6--12 & 1.91 & 2 \\
Time spent top 30\% / Time spent low 30\% & 0.80 & 1.19 \\
Money spent top 30\% / Money spent low 30\% & 1.42 & 2.67 \\\hline
\end{tabular}
\caption{Targets in the data and in the model\label{targets-main}}
\end{center}
\end{table}

\subsection{Counterfactual Simulations}

This part does two things. First, to understand the anatomy of result, I simulate individual impact of fertility shocks, comparing several counterfactual scenarios for simulated group of individuals. Second, I look at how changes in parameters driving fertility shocks and bargaining affect the observable (aggregate) moments.

\subsubsection{Individual Impact}

This part looks at counterfactual impact of the fertility shocks. To do this, I generate two sequences of shocks from the estimated distribution, $\mathcal{S}_{m,p}$ and $\mathcal{S}_{m,np}$. These sequences are identical in terms of all shocks and in both of them all females meet the first partner at $\bar{t}$, but in $\mathcal{S}_{m,p}$ they get an unexpected fertility shock at $\bar{t}$ and in $\mathcal{S}_{m,np}$ they do not. After $\bar{t}$ the sequences are identical (including chances of meeting other partners if the first marriage does not happen/breaks up).  For an auxiliary considerations I also generate sequence of shocks $\mathcal{S}_{nm}$, where at $\bar{t}$ (and before) the agents did not meet anyone.

First, one can look at average impact of the shock, presented at Figure \ref{individual-av}. I take $\bar{t} = 25$. On the top panel, I present probabilities of being married for those who agreed to marry to the partner they met at 25 (so at 25 this is one). As we can see, In the middle, there are unconditional shares of married for two groups. In the bottom I present average utility surplus (love shock) of married couples.

There are two important observations. First, conditional on having satisfactory partner women who did receive fertility shocks have around higher chances of not being married (to anyone) in 10 years. Second, threshold for satisfactory partners falls upon receiving fertility shock: more people marry if the shock happens and average surplus of married couples is lower. This indicates the presence of females who married \emph{because} the pregnancy shock happened (meaning that they would reject the partner if they were not pregnant). Then we can notice that after a number of years surviving couples become very identical in the amount of ``love shock'', meaning that couples with low $\psi$ break up. Third, in the model the largest impact of fertility shock is hurting future marriage prospective for those who stayed a single mother. This is curious although pretty speculative, as how the lifecycle of single mothers is modeled in pretty stylized way.

In the second part, I decompose the observable impact on three groups: those who would marry the partners they met with and without fertility shocks (always-marry), those who marry only if fertility shocks happen (compliers) and those who never marry their matches (never-marry). In principle, there could exist a group of those who marry only without the shock (defiers), but it is empty in my simulations. Figure \ref{ind-decomp} shows this decomposition. Two dashed lines correspond to the top panel of Figure \ref{individual-av}. The average impact of shock (dashed line with crosses) is a weighted average of two groups (presented by two solid lines). We can see that mainly the divorce probability is driven by small (5.5\%) group of compliers who have dramatically lower probabilities to stay married. Effect on fertility shocks on those who would've married anyway still persists, but is around 2-3\% (this is, however, substantial if we compare it to average probability of not-staying-married without shock that is slightly above 10\%.

In terms of characteristics, the group of compliers have about 14\% higher median savings and 1.7\% higher median labor income relative to the general population. What makes them compliers is mainly the value of love shock $\psi$: for this group, it is lower than the shock for general population of new couples by about 2/3 of initial standard deviation (i.e. $\sigma(\psi_0)$). This sort of highlights that  the compliers are generally more picky group (that is roughly consistent with evidence on education/income partition) who got a pregnancy shock together with unlucky draw of $\psi$. This is however partially hardcoded in the model, as I target different divorce probabilities for higher-income women.

\begin{figure}[h!]
\centering
\includegraphics[width=0.85\linewidth]{individual-av.eps}
\caption{Average impact of fertility shock at $t=25$\label{individual-av}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.85\linewidth]{ind-decomp.eps}
\vspace{-4em}
\caption{Decomposition of the average impact\label{ind-decomp}}
\end{figure}

\subsubsection{Aggregate Impact}
I consider two counterfactual scenarios: when unexpected fertility never happens at the moment people meet and when it never happens at all (but people still can make kids by choice). I compare key targets in Table \ref{cf-shocks}. First, absence of fertility shocks dramatically reduces early fertility (in the model every couple before 37 can have a baby by choice). Then, fertility shocks at meeting are large driver of observable divorce rates, especially for higher-income couples. Finally, fertility shocks in couples (i.e. mistimed births) are also large contributor to the divorce rate,disabling them eliminates early divorces almost completely (though is not so important for later divorces relative to shocks at births). 
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|}\hline
Target & Baseline & No shocks at meeting & No shocks at all \\\hline
\multicolumn{4}{|c|}{Marriage Dimension} \\\hline
Never married by 25  & $67.8\%$ & $68.2\%$ & $66.4\%$  \\
Never married by 30 & $25.2\%$ & $21.8\%$ & $21.4\%$  \\
Never married by 40 & $20.6\%$ & $18.1\%$ & $17.5\%$  \\\hline
\multicolumn{4}{|c|}{Fertility Dimension} \\\hline
No kids at 25 & $59.0\%$ & $85.8\%$ & $99.8\%$ \\
No kids at 30 (low $30\%$ income) & $76.0\%$  &   $53.5\%$ & $92.8\%$ \\
No kids at 30 (top $30\%$ income) & $70.5\%$  &   $61.9\%$ & $89.5\%$ \\
No kids at 35  & $32.9\%$  &   $26.8\%$ & $29.0\%$ \\
No kids at 40 & $19.4\%$   &  $23.2\%$  & $25.1\%$ \\\hline
\multicolumn{4}{|c|}{Divorce Dimension} \\\hline
Divorced by 25 & 6.6\%   &  2.9\% & 0.0\% \\
Divorced by 30 (low $30\%$ income)  & 14.5\% &     11.4\% & 1.4\% \\
Divorced by 30 (top $30\%$ income)  & 14.0\% &     6.5\%  & 0.0\% \\
Divorced by 35  & 15.8\% &    6.3\% & 4.5\% \\
Divorced by 40  & 15.5\% &    8.7\% & 7.4\% \\\hline
\end{tabular}
\caption{Contribution of fertility shocks \label{cf-shocks}}
\end{center}
\end{table}

%\section{Robustness Checks}
\section{Discussion And Conclusions}

\clearpage

\section*{Literature}
\bibliography{bib}
